## Do hosts with many listings employ different pricing strategies (e.g. greater price dispersion) than those with few listings?
```{r}
data <- read.csv("./base_dades_districtes.csv")
```

```{r}
# Step 1: Classify hosts based on listings count (lowest 25% vs middle 50% vs highest 25%)
lower_q <- quantile(data$host_listings_count, 0.25, na.rm = TRUE)
upper_q <- quantile(data$host_listings_count, 0.75, na.rm = TRUE)

# Create the classification
data$host_category <- ifelse(data$host_listings_count <= lower_q, "Low", 
                      ifelse(data$host_listings_count >= upper_q, "High", "Medium"))

# Step 2: Calculate price dispersion metrics by category
# Function to calculate statistics by group
calculate_stats <- function(group) {
  prices <- data$price[data$host_category == group]
  mean_val <- mean(prices, na.rm = TRUE)
  median_val <- median(prices, na.rm = TRUE)
  std_dev <- sd(prices, na.rm = TRUE)
  coef_var <- (std_dev / mean_val) * 100
  iqr_val <- IQR(prices, na.rm = TRUE)
  min_val <- min(prices, na.rm = TRUE)
  max_val <- max(prices, na.rm = TRUE)
  range_val <- max_val - min_val
  count_val <- length(prices)
  
  return(c(mean_val, median_val, std_dev, coef_var, iqr_val, min_val, max_val, range_val, count_val))
}

# Calculate statistics for each group
stats_low <- calculate_stats("Low")
stats_medium <- calculate_stats("Medium")
stats_high <- calculate_stats("High")

# Create results data frame
price_dispersion <- data.frame(
  host_category = c("Low", "Medium", "High"),
  mean_price = c(stats_low[1], stats_medium[1], stats_high[1]),
  median_price = c(stats_low[2], stats_medium[2], stats_high[2]),
  sd_price = c(stats_low[3], stats_medium[3], stats_high[3]),
  cv_price = c(stats_low[4], stats_medium[4], stats_high[4]),
  iqr_price = c(stats_low[5], stats_medium[5], stats_high[5]),
  min_price = c(stats_low[6], stats_medium[6], stats_high[6]),
  max_price = c(stats_low[7], stats_medium[7], stats_high[7]),
  price_range = c(stats_low[8], stats_medium[8], stats_high[8]),
  count = c(stats_low[9], stats_medium[9], stats_high[9])
)

# Display the summary
print(price_dispersion)

# Step 3: Create visualizations using basic R functions

# 1) Boxplot to compare price distribution
boxplot(price ~ host_category, data = data,
        main = "Price Distribution by Host Category",
        xlab = "Host Category", ylab = "Price",
        col = c("skyblue", "lightgreen", "coral"),
        ylim = c(0, quantile(data$price, 0.95, na.rm = TRUE)),
        las = 1)

# 3) Bar chart of coefficient of variation
barplot(price_dispersion$cv_price, 
        names.arg = price_dispersion$host_category,
        main = "Coefficient of Variation by Category",
        xlab = "Host Category", ylab = "CV (%)",
        col = c("skyblue", "lightgreen", "coral"))

```
This R code performs a comprehensive analysis of pricing strategies among Airbnb hosts, segmented by the number of properties they manage. The analysis follows a structured methodology to determine whether hosts with different portfolio sizes employ different pricing approaches.


The code begins by dividing hosts into three distinct categories based on their listing count:

- Low: Hosts in the bottom 25% by number of listings (casual hosts)
- Medium: Hosts in the middle 50% (moderately active hosts)
- High: Hosts in the top 25% (professional hosts with many listings)

This categorization uses quartiles as thresholds, creating a reasonable distribution that allows for meaningful comparisons between groups.

For each host category, the code calculates a comprehensive set of statistical measures to characterize pricing behavior:

- Central tendency metrics: Mean and median prices show the typical price point for each host category
- Dispersion metrics: Standard deviation and interquartile range (IQR) measure absolute price variability
- Relative dispersion: Coefficient of variation (CV) - calculated as (standard deviation/mean × 100) - measures price variability relative to the average price
- Range metrics: Minimum price, maximum price, and overall price range (max-min) show the boundaries of pricing strategies
- Sample size: Count of listings in each category ensures statistical validity
- The statistics are calculated using a custom function that processes each host category separately, ensuring clean and consistent measurement across groups.



The results obtained are:

| host_category | mean_price | median_price | sd_price | cv_price | iqr_price | min_price | max_price | price_range | count |
|---------------|------------|--------------|----------|----------|-----------|-----------|-----------|-------------|-------|
| Low           | 71.96      | 50.0         | 84.33    | 117.20   | 45        | 8         | 2000      | 1992        | 1595  |
| Medium        | 106.37     | 57.5         | 335.65   | 315.55   | 65        | 8         | 8000      | 7992        | 2163  |
| High          | 221.93     | 100.0        | 653.08   | 294.28   | 92        | 9         | 6000      | 5991        | 1260  |

These results strongly suggest that hosts with many listings do employ different pricing strategies with greater price dispersion than those with few listings.

The first thing that we can easily see is that there's a strong upward trend in both mean and median prices as we move from hosts with few listings to those with many listings. Hosts who fall within the higher quartile charge significantly higher. This can be explained due to the rising tendency of major holder of tourist accommodations in the city, who live exclusively from the tourist renting. They need to have higher prices to sustain its deal. This is causing rising prices for the locals, which are increasingly getting expelled from their neighbourhoods.

We can also observe a big difference in the dispersion of the prices of the listings, which at first sight might be counter-intuitive, because if they are normally sparsed throughout the city, the prices should be also normally distributed, which is not the case. We will focus on the Coefficient of Variation (CV), which is a statistical measure that describes the dispersion of data points around the mean, relative to the mean itself. For the low-owning hosts, this is 117% which means that the low-listing hosts often adopt straightforward pricing strategies. Having only one or two properties means they're less likely to experiment with diverse price points and more likely to choose a “safe” middle-ground price based on limited market research or platform suggestions. Hosts with few listings regularly prioritize consistency and predictability over maximum revenue optimization. They may set prices and rarely adjust them, while hosts with many listings actively manage and diversify their pricing. It also can be explained by their low operating scale, which without the economies of scale in operations, low-listing hosts typically have similar cost structures across their limited properties, resulting in comparable pricing requirements to achieve profitability.

The mid-listing hosts have an extremely high coefficient of variation: 315%, which is even higher than the high-listing hosts, being at 294%. This makes sense when we think about the personal situation of a mid-listing host. Medium hosts are often in a transitional phase between casual and professional hosting. This middle stage typically involves active experimentation, which causes a higher range of prices, target demographics and seasonal changes. This phase should be got over with for the high-listing hosts, which given the experience, they are in a more standardized phase with more market domain and better predictability, that provoke a relatively-lower price range caused by some expenses being cut off which reflect in the final price. 

This different boundaries also reflect when we see the price range difference, being higher in the mid-owning hosts, explained by this experimental phase we've already talked about. We could now try to compare the proportion of different types of listing between the high range owners and the mid-ones:
```{r}

room_type_by_category <- table(data$host_category, data$room_type)
room_type_prop <- prop.table(room_type_by_category, 1) * 100  # Percentage by row

print("Distribution of the listings categories by host type(%):")
print(round(room_type_prop, 2))

# Create stacked barplot from the room type proportions
barplot(t(room_type_prop), 
        main = "Accommodation Types by Host Category (%)",
        xlab = "Host Category", 
        ylab = "Percentage (%)",
        col = c("steelblue", "coral", "lightgreen"),
        legend.text = colnames(room_type_prop),
        args.legend = list(x = "topright", cex = 0.8))  
```
These results confirm our theory, where we can see that mid-listing hosts tend to have a much more distributed portfolio than low-listing hosts, which are more focused on private room hosting. The high-listing hosts are more focused on entire apartment hosting. The mid-listing hosts fall in the middle, which corroborates this experimenting phase.