##  Do hosts with fast response times capture more last-minute bookings (low minimum_nights & low availability) and thereby higher average occupancy/review rates?

```{r}
dataset <- read.csv("./base_dades_districtes.csv")
```

```{r}
# Step 1: Classify response times as fast or slow
# Fast response is defined as "within an hour" OR "within a few hours"
dataset$response_speed <- ifelse(dataset$host_response_time %in% c("within an hour","within a few hours"), 
                                "Fast", 
                                ifelse(dataset$host_response_time == "N/A", 
                                       NA, "Slow"))

# Step 2: Define last-minute friendly listings (low minimum nights & low availability), medians are the threshold
min_nights_threshold <- median(dataset$minimum_nights, na.rm = TRUE)
availability_threshold <- median(dataset$availability_365, na.rm = TRUE)

dataset$last_minute_friendly <- ifelse(dataset$minimum_nights <= min_nights_threshold & 
                                      dataset$availability_365 <= availability_threshold,
                                     "Last-minute friendly", "Regular")

# Step 3: Create contingency table to see relationship between response time and last-minute friendliness
# Remove NA values for clean analysis
response_data <- dataset[!is.na(dataset$response_speed), ]
response_lastminute_table <- table(response_data$response_speed, response_data$last_minute_friendly)
response_lastminute_prop <- prop.table(response_lastminute_table, 1) * 100  # Row percentages

# Display results
print("Percentage of last-minute friendly listings by response time:")
print(round(response_lastminute_prop, 2))

# Step 4: Calculate occupancy and rating statistics by group
# Using number_of_reviews as a proxy for occupancy (more reviews = higher occupancy)
occupancy_by_response <- tapply(response_data$number_of_reviews, response_data$response_speed, 
                               function(x) c(mean = mean(x, na.rm = TRUE),
                                           median = median(x, na.rm = TRUE)))

ratings_by_response <- tapply(response_data$review_scores_rating, response_data$response_speed, 
                             function(x) c(mean = mean(x, na.rm = TRUE),
                                          median = median(x, na.rm = TRUE)))

print("Occupancy statistics (number of reviews) by response time:")
print(occupancy_by_response)

print("Rating statistics by response time:")
print(ratings_by_response)

# Step 5: Visualization with simple plots
# Bar chart showing proportion of last-minute friendly listings
barplot(response_lastminute_prop[, "Last-minute friendly"], 
        main = "% of Last-minute Friendly Listings by Response Time",
        xlab = "Host Response Time", 
        ylab = "Percentage (%)",
        col = c("coral", "steelblue"),
        names.arg = c("Fast", "Slow"))

# Box plots comparing reviews and ratings
par(mfrow = c(1, 2))
boxplot(number_of_reviews ~ response_speed, data = response_data,
        main = "Number of Reviews by Response Time",
        xlab = "Host Response Time",
        ylab = "Number of Reviews",
        col = c("coral", "steelblue"))

boxplot(review_scores_rating ~ response_speed, data = response_data,
        main = "Ratings by Response Time",
        xlab = "Host Response Time",
        ylab = "Rating Score",
        col = c("coral", "steelblue"))
par(mfrow = c(1, 1))
```
This code defines as fast responders those who have response times of either "within an hour" or "within a few hours". By having definition of fast and slow responders, we can analyze if there's a meaningful threshold effect between hosts who respond promptly (within hours) versus those who take longer (days or more). Then we define the last-minute friendly listings, which will be those who have two key characteristics:

1.  fewer than the median number of minimum nights - indicating flexibility for short stays
2. Show less than median availability throughout the year - suggesting they're already successful at securing bookings

This definition of "last-minute friendly" captures properties that are structurally more accommodating to spontaneous travelers who might be looking for quick, short-term arrangements. The last-minute friendly listings will be those who are under both thershold.

After creating these two groups, we check if hosts who respond quickly have more listings that are good for last-minute bookings. We also look at two important things:

- Do fast-responding hosts get more bookings? (We count reviews to estimate this)
- Do guests like staying with fast responders better? (We look at their rating scores)

This helps us understand if being quick to answer messages helps hosts get more last-minute bookings and happier guests.

At first, we'd determined that the fast responders would be those who answer within an hour, whom results were:

       Last-minute friendly Regular
  Fast                29.78   70.22
  Slow                28.60   71.40

As we can see, there's is not much difference between the fast and slow answering hosts, suggesting that responding speed doesn't influence how much last-minute friendly is a listing. So we decided to take a less radical position and accept the within a few hours hosts as fast responders to see if anything changes. The new results were:

       Last-minute friendly Regular
  Fast                27.13   72.87
  Slow                35.42   64.58
  
These results now suggest that there is actually an inverse relationship - slower responders appear to have a higher percentage of last-minute friendly listings (35.42% vs 27.13%). This is counter-intuitive and might indicate that response time alone isn't the determining factor for last-minute bookings. Perhaps hosts who take longer to respond have other attractive features or competitive pricing that make their properties appealing despite the slower response.

The other graphs reveal an interesting relationship between host response time and success metrics on Airbnb. Fast responders clearly attract more bookings, shown by their significantly higher review counts, with some popular listings receiving over 400 reviews. However, quick responses don't automatically translate to better guest experiences. Both fast and slow responders maintain high ratings (mostly above 90), but slow responders actually show more consistent top scores, while some fast responders have concerning ratings below 80. Most surprisingly, our analysis found that slow responders have more "last-minute friendly" listings (35.42%) than fast responders (27.13%), challenging our initial assumption. This suggests that while quick communication helps secure bookings generally, other factors like location, pricing, and booking flexibility may matter more to last-minute travelers than simply how fast a host responds to messages. The data indicates that success in the touristic sector in Barcelona comes from balancing responsiveness with overall quality and guest experience, rather than just focusing on response speed.